---
title: "LBB1"
author: "Luthfi"
date: "2023-09-23"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
    df_print: paged
---

# 1. REAL ESTATE MELBOURNE

This is a snapshot of a dataset created by Tony Pino. link: <https://www.kaggle.com/datasets/dansbecker/melbourne-housing-snapshot>

It was scraped from publicly available results posted every week from Domain.com.au. He cleaned it well, and now it's up to you to make data analysis magic. The dataset includes Address, Type of Real estate, Suburb, Method of Selling, Rooms, Price, Real Estate Agent, Date of Sale and distance from C.B.D.

## Notes on Specific Variables

Rooms: Number of rooms

Price: Price in dollars

Method: S - property sold; SP - property sold prior; PI - property passed in; PN - sold prior not disclosed; SN - sold not disclosed; NB - no bid; VB - vendor bid; W - withdrawn prior to auction; SA - sold after auction; SS - sold after auction price not disclosed. N/A - price or highest bid not available.

Type: br - bedroom(s); h - house,cottage,villa, semi,terrace; u - unit, duplex; t - townhouse; dev site - development site; o res - other residential.

SellerG: Real Estate Agent

Date: Date sold

Distance: Distance from CBD

Regionname: General Region (West, North West, North, North east ...etc)

Propertycount: Number of properties that exist in the suburb.

Bedroom2 : Scraped \# of Bedrooms (from different source)

Bathroom: Number of Bathrooms

Car: Number of carspots

Landsize: Land Size

BuildingArea: Building Size

CouncilArea: Governing council for the area

---------
The First step is inserting the csv file into R located in data_input and then installing the necessary plugins including dplyr, ggplot2, and leaflet that will be used.

```{r}
#Read data csv
melb <- read.csv("data_input/melb_data.csv")

library(dplyr) #install package dplyr to manipulate character data
library(ggplot2) #install package ggplot2 for data visualization
library(leaflet) #install package leaflet for map plotting

```

#2. DATA PREPERATION

Next, we will inspect the 'melb' folder and then remove duplicate rows based on the 'Address' column. Additionally, we will eliminate any unnecessary columns.

```{r}
#First, we read the .csv file and use only the relevant columns
melb
#Clean duplicates based on address
melb <- melb[!duplicated(melb$Address),]
# Removing unnecessary columns using select()
melb2 <- melb[,!names(melb) %in% c("Address", "Postcode", "Propertycount", "CouncilArea", "YearBuilt", "BuildingArea")]
# Removing NA Rows and "0"
melb2 <- melb2[melb2$Landsize != 0, ]
```

After removing the unnecessary rows and columns, we will proceed to convert the data types to facilitate plotting. The firt column we convert is melb$Date using the "%d/%m/%Y", and the rest as Factor and integer.

```{r}
# Converting Date datatype
#Converting Date
melb2$Date <- as.Date(melb2$Date, format = "%d/%m/%Y") 

#Converting as Factor & Integer
mutate(.data = melb2,
       Suburb = as.factor(Suburb),
       Method = as.factor(Method),
       Type = as.factor(Type),
       Regionname = as.factor(Regionname),
       Bedroom2 = as.integer(Bedroom2),
       Bathroom = as.integer(Bathroom),
       Car = as.integer(Car),
       Landsize = as.integer(Landsize)
       )
       


```
Next, we will rename the categories in the 'Regionname' and 'Type' columns to make them easier to read and eliminate any unnecessary duplicates.

```{r}
#Renaming Categories for Regionname
melb2 <- melb2 %>%
  mutate(Regionname = recode(Regionname, "Northern Victoria" = "Northern Metropolitan"), 
        Regionname = recode(Regionname, "Eastern Victoria" = "Eastern Metropolitan"),
        Regionname = recode(Regionname, "Western Victoria" = "Western Metropolitan"))

#Renaming Categories for Type
melb2 <- melb2 %>%
  mutate(Type = recode(Type, "h" = "Housing"), 
        Type = recode(Type, "u" = "Unit"),
        Type = recode(Type, "t" = "Townhouse"))
```

#3. DATA EXPLORATION

Calculate the price per square meter for each property in the dataset and then removes any outliers (values outside the 1.5*IQR range) from the 'PricePerSQM' column, resulting in a dataset with potentially more reliable data for analysis or visualization.

```{r}
# Creating Price Per Squaremeter Column
melb2$PricePerSQM <- (melb2$Price / melb2$Landsize)

#Removing outlier
Q1 <- quantile(melb2$PricePerSQM, 0.25)
Q3 <- quantile(melb2$PricePerSQM, 0.75)
IQR <- Q3 - Q1
LB <- Q1 - 1.5*IQR
UB <- Q3 + 1.5*IQR

melb2 <- subset(melb2, PricePerSQM >= LB & PricePerSQM <= UB)
```

According to the histogram below, the most frequent number of rooms is 3.

```{r}
#Number of Rooms in Building
hist(melb2$Rooms, col = c('lightblue'), main = "Number of Rooms", xlab = "Rooms", density = 50)
#mean(melb2$Rooms) # Center of Data = 2.977419 (Rooms)

```

Based on the histogram below, the most frequently occurring distance to the CBD falls within the range of 5 to 15, with a median distance of 10.1.

```{r}
#Distance
hist(melb2$Distance, col = c('orange'), main = "Melbourne Housing Distance", xlab = "Distance", density = 50)
#median(melb2$Distance) # Center of Data = 10.1 (Distance)
```

According to the table below, the suburb with the highest count of property sold is Reservoir with a total 160 properties.

```{r}
#Top 10 suburb with the highest number of sold properties

freq_suburb <- table(melb2$Suburb) # Create frequency table
freq_suburb_df <- as.data.frame(freq_suburb) # Create data frame
colnames(freq_suburb_df) <- c("Suburb", "PropertiesSold") # Renaming columns
freq_suburb_df <- freq_suburb_df[order(-freq_suburb_df$PropertiesSold),]
top10_freq_suburb <- freq_suburb_df[1:10, ]
top10_freq_suburb
```

According to the table below, the suburb with the highest average price per SQM is Princes Hill with a price of $7453.

```{r}
# Top 10 Suburb with highest average of buying price per square meter
average_price <- melb2 %>%
  group_by(Suburb) %>%
  summarise(Average_Price_Per_Square_Meter = median(PricePerSQM, na.rm = TRUE)) # Creating a table
average_price <- average_price[order(-average_price$Average_Price_Per_Square_Meter),] # Sorting order
colnames(average_price) <- c("Suburb", "Average price Per SQM") # Renaming columns
top10_freq_suburb2 <- average_price[1:10, ]
top10_freq_suburb2

```

The Boxplot below indicates that the type of housing that is the most frequently sold is a townhouse type, and then housing, with a close third unit.

```{r}
boxplot(melb2$PricePerSQM~ melb2$Type,horizontal = F, col = c('orange'), main = "Melbourne Housing Price & Type", xlab = "Type", ylab = "Price Per SQM")
# h - house,cottage,villa, semi,terrace
# u - unit, duplex
# t - townhouse; dev site - development site
```

## Relations between Price and Distance
The plot below indicates a correlation between the distance of a house and the price per SQM. The closer you are to the CBD, the higher the price. Additionally, the plot also reveals that the lowest price located near the CBD is a unit type and the most frequent one is a housing type.

```{r}
#Plotting Price PerSQM and Distance to CBD
#cor(melb2$Distance, melb2$PricePerSQM) = -0.4333912
ggplot(data = melb2, aes(x = Distance, y = PricePerSQM, color = Type)) +
  geom_point() +
  labs(
    title = "Relations between Price and Distance", subtitle = "Melbourne Housing",
    x = "Distance",
    y = "Price per SQM"
  ) 
```
## Melbourne Housing Rooms & Price
From the plot below, it is found that housing units with two rooms have the highest median price per square meter."
```{r}
#Plotting Price PerSQM and number of rooms
boxplot(melb2$PricePerSQM~ melb2$Rooms,horizontal = F, col = c('orange'), main = "Melbourne Housing Rooms & Price", xlab = "Rooms", ylab = "Price Per SQM")
```
## Melbourne Housing Bedroom & Price
From the plot below, it is found that housing units with two bedrooms have the highest median price per square meter.

```{r}
#Plotting Price PerSQM and number of bedroom
boxplot(melb2$PricePerSQM~ melb2$Bedroom2,horizontal = F, col = c('orange'), main = "Melbourne Housing Bedroom & Price", xlab = "Bedroom", ylab = "Price Per SQM")
```
## Melbourne Housing Bathroom & Price
From the plot below, it is found that housing units with four bathrooms have the highest median price per square meter.

```{r}
#Plotting Price PerSQM and number of bathroom
boxplot(melb2$PricePerSQM~ melb2$Bathroom,horizontal = F, col = c('orange'), main = "Melbourne Housing Bathroom & Price", xlab = "Bathroom", ylab = "Price Per SQM")
```



## Melbourne Housing Region & Price
According to the barplot below, it is observed that housing units located in the Southern Metropolitan region have the highest average price compared to other regions, while the South-Eastern region has the lowest average price. Typically, the housing type with the highest cost is 'Housing,' followed by 'Townhouse,' and then 'Unit'.

```{r}
#Plotting Price PerSQM and region

ggplot(data = melb2, aes(x = Regionname, y = PricePerSQM, color = Type)) +
  geom_col(mapping = aes(fill=Type), position = "dodge") +
   theme(
    text = element_text(size = 7),# Adjust the font size
    panel.background = element_rect(fill="white")
  )
```

Below is a map showing the housing sold in Northern Metropolitan region in Melbourne.


```{r}
# Map
melbmap <- subset(melb2, melb2$Regionname == "Northern Metropolitan")
melbmap <- data.frame(lat=melbmap$Lattitude, # latitude
                   lng=melbmap$Longtitude) # longitude
set.seed(50)
melbmap <- melbmap[sample(nrow(melbmap), 25), ]
# get icon
ico <- makeIcon(
    iconUrl = "https://cdn-icons-png.flaticon.com/512/25/25694.png",
    iconWidth=41/2,
    iconHeight=41/2
)


# create a leaflet map widget
map <- leaflet()
# add tiles from open street map
map <- addTiles(map = map)
# add markers
addMarkers(map = map, data = melbmap, icon = ico)

```


#4.  SUMMARY

📌 
In Melbourne, some of the most expensive houses often feature two roomy bedrooms and an impressive four bathrooms. Princes Hill, located in the Northern Metropolitan area, is one place where you can find such luxurious properties. 

------------------------------------------------------------------------

DOT POINTS

1. The closer you are to the CBD, the higher the price

2.  The most sold properties is located at Reservoir Suburb

3.  Highest housing price in general

- Highest price of housing is located at Princes Hill which is part of Northern Metropolitan
- Highest price of housing type is a townhouse
- 2 Rooms
- 2 Bedroom
- 4 Bathroom

